Navigating Trends: Analyzing Ticket Sales and Redemption Patterns For Toronto Island Ferries

### 2024-09-24 Paper/Paper.qmd
added finishing touches, fixed cross referencing.

A sample of the cleaned data set can be found in table 1 [@tbl-datapreviewtable] showcasing the variables and the structure of the data.

```{r}

# CLEANED DATA PREVIEW TABLE

#| label: tbl-datapreviewtable
#| tbl-cap: "Sample of cleaned Ferry Ticket Counts Data"
#| warning: false
#| echo: false
#| message: false

@@

kable(data_preview,
      caption = "Sample of Cleaned Ferry Ticket Counts Data",
      col.names = c("ID", "Time", "Tickets Redeemed", "Tickets Sold", "Date"), 
    digits = 1,
    booktabs = TRUE,
    linesep = "", 
    align = c("c", "c", "c", "c", "c")
  )

@@

The summary statistics are provided in table 2 [@tbl-summary] and allow us to get an understanding of the data in a very comprehensible manner. The high volume of tickets sold and redeemed indicates a healthy and significant demand of the ferry and the closeness of the values conveys strong customer engagement and a reliable service. On average, 46.92 tickets were sold every 15 minutes during working hours, and 47.37 tickets were redeemed, indicating that most customers utilized their tickets. However, the small gap implies some no-shows or unused tickets. The maximum numbers of redemption and sales tickets (7,216 and 7,229) showcases the ferry's ability to handle high customer volumes and the closeness of these values tells us that, on busy days, the ferry usually operates at capacity.

@@

```{r}
# SUMMARY TABLE

#| label: tbl-summary
#| tbl-cap: "Summary Statistics of Ticket Counts"
#| warning: false
#| echo: false
#| message: false

@@

# Print the summary table 
kable(summary_table, caption = "Summary Statistics of Ticket Counts", format = "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, position = "center", (latex_options = c("hold_position", "repeat_header"))) %>%  
  column_spec(1:5, width = "2cm")

@@

After studying the summary statistics, we have an idea of the capacity and performance of the ferry, along with the relationships between the variables. Next, with the use fo figure 1 [@fig-timeseriesanalysis], we can study the patterns of ticket counts from 2018 to 2024. For figure1, we can see that there is a definitive pattern in the usage numbers when looking at each year. We see that there are spikes during the mid-year months for each year, indicating higher popularity of the Toronto Islands during those times and suggesting that the islands serve as a significant recreational destination.

We can also see that for every year, there is a slow climb from the beginning of the year (which starts very low) indicating an off-season and the start of the tourist season. Conversely, there is a gradual decline during the end of the year (also indicating an off-season) which confirms a seasonal drop in demand as the cold weather deters visitors. The highs in each year are steady at 800 - 1000 for the number of tickets sold in one interval, however, the highs for the number of tickets redeemed are usually lower, but ever-fluctuating. This fluctuation and discrepency could indicate that not all passengers redeem their tickets within the same time frame or that there are differences in purchasing behavior.

Notably, figure 1 [@fig-timeseriesanalysis] also shows an anomaly during the years of 2020 and 2021, conveyed by highly depressed sales and redemption ticket counts. This can be attributed to the COVID-19 pandemic, which lead to widespread public facility closures and reduced operational capacities. The peak ticket count in 2020, reaching only about 200, is in high contrast with every other year, highlighting the pandemic's massive impact on public transportation and tourism.

```{r}

#TIME SERIES ANALYSIS PLOT

#| label: fig-timeseriesanalysis
#| fig-cap: "Tickets Sold and Redeemed by Date"
#| warning: false
#| echo: false
#| message: false
#| fig-align: "center"

@@

# Histogram for Tickets Sold and Redeemed
ggplot(long_ticket_data, aes(x = Date, y = Count, fill = Ticket_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
      caption = "Figure 1:Tickets Sold and Redeemed by Date",
       x = "Date",
       y = "Total Count") +
  scale_fill_manual(values = c("Tickets_Sold" = "blue", "Tickets_Redeemed" = "orange"), 
                    labels = c("Tickets Sold", "Tickets Redeemed")) +
  ylim(0, 1250) +
  theme_minimal()

@@

Figure 2 [@fig-monthanalysis], shows a monthly analysis of redemption and sales ticket counts for each year. The data in this figure has been grouped together and summed. i.e. the ticket counts for each of the 12 months in every year from 2018 to 2024 has been summed and aggregated so that we can see the monthly trends over the past 6 years.

@@

```{r}

## MONTH BY MONTH ANALYSIS

#| label: fig-monthanalysis
#| fig-cap: "Aggregate Monthly Ticket Counts (Sold and Redeemed)"
#| warning: false
#| echo: false
#| message: false
#| fig-align: center

@@

  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.7) +
  labs(
    x = "Month", 
    y = "Total Count", 
    caption = "Figure 2: Aggregate Monthly Ticket Counts (Sold and Redeemed)", fill = NULL) +

>>>>

A sample of the cleaned data set can be found in [@tbl-datapreviewtable] showcasing the variables and the structure of the data.

```{r}
#| label: tbl-datapreviewtable
#| tbl-cap: "Sample of cleaned Ferry Ticket Counts Data"
#| warning: false
#| echo: false
#| message: false

@@

kable(data_preview,
      col.names = c("ID", "Time", "Tickets Redeemed", "Tickets Sold", "Date"), 
    digits = 1,
    booktabs = TRUE,
    linesep = "", 
    align = c("c", "c", "c", "c", "c")
  )

@@

The summary statistics are provided in [@tbl-summary] and allow us to get an understanding of the data in a very comprehensible manner. The high volume of tickets sold and redeemed indicates a healthy and significant demand of the ferry and the closeness of the values conveys strong customer engagement and a reliable service. On average, 46.92 tickets were sold every 15 minutes during working hours, and 47.37 tickets were redeemed, indicating that most customers utilized their tickets. However, the small gap implies some no-shows or unused tickets. The maximum numbers of redemption and sales tickets (7,216 and 7,229) showcases the ferry's ability to handle high customer volumes and the closeness of these values tells us that, on busy days, the ferry usually operates at capacity.

@@

```{r}
#| label: tbl-summary
#| tbl-cap: "Summary Statistics of Ticket Counts"
#| warning: false
#| echo: false
#| message: false

@@

# Print the summary table 
kable(summary_table, format = "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, position = "center", (latex_options = c("hold_position", "repeat_header"))) %>%  
  column_spec(1:5, width = "2cm")

@@

After studying the summary statistics, we have an idea of the capacity and performance of the ferry, along with the relationships between the variables. Next, with the use of[@fig-timeseriesanalysis], we can study the patterns of ticket counts from 2018 to 2024. For figure 1, we can see that there is a definitive pattern in the usage numbers when looking at each year. We see that there are spikes during the mid-year months for each year, indicating higher popularity of the Toronto Islands during those times and suggesting that the islands serve as a significant recreational destination.

We can also see that for every year, there is a slow climb from the beginning of the year (which starts very low) indicating an off-season and the start of the tourist season. Conversely, there is a gradual decline during the end of the year (also indicating an off-season) which confirms a seasonal drop in demand as the cold weather deters visitors. The highs in each year are steady at 800 - 1000 for the number of tickets sold in one interval, however, the highs for the number of tickets redeemed are usually lower, but ever-fluctuating. This fluctuation and discrepancy could indicate that not all passengers redeem their tickets within the same time frame or that there are differences in purchasing behavior.

Notably, [@fig-timeseriesanalysis] also shows an anomaly during the years of 2020 and 2021, conveyed by highly depressed sales and redemption ticket counts. This can be attributed to the COVID-19 pandemic, which lead to widespread public facility closures and reduced operational capacities. The peak ticket count in 2020, reaching only about 200, is in high contrast with every other year, highlighting the pandemic's massive impact on public transportation and tourism.

```{r}
#| label: fig-timeseriesanalysis
#| fig-cap: "Tickets Sold and Redeemed per 15 min Interval"
#| warning: false
#| echo: false
#| message: false
#| fig-align: center

@@

# Histogram for Tickets Sold and Redeemed
ggplot(long_ticket_data, aes(x = Date, y = Count, fill = Ticket_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Date",
       y = "Total Count") +
  scale_fill_manual(values = c("Tickets_Sold" = "blue", "Tickets_Redeemed" = "orange"), 
                    labels = c("Tickets Sold", "Tickets Redeemed")) +
  ylim(0, 1250) +
  theme_minimal() + 
  theme(axis.title.x = element_text(margin = margin(t = 10)), 
        axis.title.y = element_text(margin = margin(r = 10)))

@@

[@fig-monthanalysis] shows a monthly analysis of redemption and sales ticket counts for each year. The data in this figure has been grouped together and summed. i.e. the ticket counts for each of the 12 months in every year from 2018 to 2024 has been summed and aggregated so that we can see the monthly trends over the past 6 years.

@@

```{r}
#| label: fig-monthanalysis
#| fig-cap: "Aggregate Monthly Ticket Counts (Sold and Redeemed)"
#| warning: false
#| echo: false
#| message: false
#| fig-align: center

@@

  geom_bar(stat = "identity", position = position_dodge(width = 0.9), alpha = 0.7) +
  labs(
    x = "Month", 
    y = "Total Count", 
     fill = NULL) +
